---
title: "Lab 1"
author: "Conie O'Malley"
date: "`r Sys.Date()`"
format: 
  html:
    self-contained: true
    code-fold: true
    code-tools: true
execute:
  echo: false
  freeze: false
editor: visual
---

```{r setup code chunk}
# libraries - commented out
library(reticulate)
library(tidyverse)
library(ggplot2)
#use_condaenv("datascience", required = FALSE) # set my environment
```

## Part 1: Hello World and Beyond

### Issuing Interactive Commands & Adding Comments

```{r}
print("Hello World!") # first line r of code
```

```{r}
print("Hello Stats!") # second line r of code
```

```{python}
print("Hello World!") # frist line of python code
print("Hello Stats!") # second line of python code
```

### Doing Simple Math Calculations

```{r}
1 + 2 + 3 + 4 + 5 
sum(1:5) # alternate way to write the code
```

### Creating and Using Vectors and Operations

```{r}
c(1, 2, 3, 4, 5) # concatenate 
1:5 # sequence operator
sum(1:5) # addition using the sequence operator
```

### Storing and Calculating Values

```{r}
x <- 1:5 # vector assignment
y <- 10 # vector assignment
x+y # vector addition
z <- x+y # vector assignment
z
h <- "Hello" # vector assignment
h
hw <- c("Hello", "World!") # vector concatenation 
print(hw) # print view
paste(hw) # paste view
```

### Navigating the RStudio Workspace

```{r}
ls() # view created objects
rm("z") # remove "z" object
ls() # confirm removal
```

### More Practice Vectorizing & Vectors of Unequal Length

```{r}
baskets.of.granny <- c(12, 4, 4, 6, 9, 3)
sum(baskets.of.granny)

firstnames <- c("John", "Jacqueline", "Robert") # create vector list
lastname <- "Kennedy" # create vector
paste(firstnames, lastname) # paste vectors

lastnames <- c("Kennedy", "Kennedy-Onnasis") # create vector list
paste(firstnames, lastnames) # paste vectors
```

## Part 2: Statistical Analysis with R

### Scatter Plot

```{r}
ggplot(mpg, aes(displ, hwy)) + # create a scatter plot
  geom_point() +
  labs(title = "Displacement vs. Highway MPG Scatterplot", x = "Displacement (litres)", y = "Highway MPG") + # apply labels
  theme(plot.title = element_text(hjust = 0.5))
```

#### Analysis

There is a inverse relationship between the volume of engine displacement and highway miles per gallon. The relationship appears to be slightly curvilinear, but cannot be confirmed without a residuals plot.

### Box Plot

```{r}
ggplot(mpg, aes(class, hwy)) + # create a box plot
  geom_boxplot() +
  labs(title = "Class vs. Highway MPG Box Plot", x = "Class", y = "Highway MPG") + # apply labels
  theme(plot.title = element_text(hjust = 0.5))
```

#### Analysis

There seems to be some commonality among the types *2seater*, *compact*, *midsize*, and *subcompact*, because they all have overlapping box plot ranges. *pickup* and *suv* also fall into their own common area because of overlapping ranges, while the *minivan* class has no overlaps. There are a considerable amount of outliers in the *suv* class - meaning that, at first glance, this is the most varied class of vehicles, where as *2seater* has no outliers and a very tight range - meaning that, at first glance, this is likely the most homogeneous class of vehicles.

### Computing Basic Statistics

```{r}
mean(economics$unemploy) # calculate mean
var(economics$unemploy) # calculate variance
sd(economics$unemploy) # calculate standard deviation
min(economics$unemploy) # calculate min
max(economics$unemploy) # calculate max
median(economics$unemploy) # calculate median
cor(economics$pce, economics$psavert) # calculate correlation
pce_psavert_cor <- round(cor(economics$pce, economics$psavert),4)
```

#### Analysis

There is a strong, negative correlation between *pce* and *psavert* (`r pce_psavert_cor`).
