---
title: "Lab 8"
author: "Conie O'Malley"
date: "`r Sys.Date()`"
format: 
  pdf: 
    titlepage: true
    toc: true
    toc-depth: 2
    code-block-wrap: true
    number-sections: true
execute:
  echo: true
  freeze: false
  error: false
  jupyter: python3
python:
  version: /Users/coniecakes/anaconda3/envs/datascience/bin/python
---

# Pre Lab
```{r libraries}

required_packages <- c("tidyverse", "quanteda", "readtext", "stm", "stminsights", "wordcloud", "gsl", "topicmodels", 
                        "caret", "gutenbergr", "tidytext", "quanteda.textmodels", "tm", "igraph", "ggraph", "widyr", 
                        "jsonlite", "factoextra", "janeaustenr", "cluster", "SnowballC", "proxy", "stringr", "textclean",
                        "dendextend", "ggdendro", "plotly", "reticulate")

```

## Deliverable 1: Get your working directory and paste below:

"/Users/coniecakes/Library/CloudStorage/OneDrive-Personal/001. Documents - Main/023. Programming Tools/R Studio/AdvTxtAnalytics"

# Part 1: Introduction to K-Means Clustering with Synthetic Data

```{r}

texts <- c(
    # Sports articles
    "The football team scored three touchdowns. The quarterback threw perfect passes.",
    "The basketball team won with slam dunks and three-point shots.",
    "The baseball pitcher threw a perfect game with many strikeouts.",
    "The soccer team scored two goals and won the championship.",
    # Technology articles
    "The computer system processed data using advanced algorithms and databases.",
    "The network router managed bandwidth through secure encryption protocols.",
    "The software code executed functions through compiled programming syntax.",
    "The digital platform integrated APIs with cloud computing architecture.",
    # Food articles
    "The professional chef created sauces and prepared gourmet dishes.",
    "The master baker produced artisan breads and pastries daily.",
    "The culinary team seasoned meats and roasted vegetables.",
    "The kitchen staff garnished plates and plated entrees."
)

```

## Add category labels

```{r categories}

categories <- rep(c("Sports", "Technology", "Food"), each = 4)
cat("Initial document count:", {length(texts)})

```

# Create corpus and preprocess text

```{r corpus}

corpus <- tm::Corpus(tm::VectorSource(texts))
cat("Corpus size:", {length(corpus)})

cleanCorpus <- function(corpus) {
    corpus <- tm::tm_map(corpus, tm::content_transformer(tolower))
    corpus <- tm::tm_map(corpus, tm::removePunctuation)
    corpus <- tm::tm_map(corpus, tm::removeNumbers)
    corpus <- tm::tm_map(corpus, tm::removeWords, tm::stopwords("english"))
    corpus <- tm::tm_map(corpus, tm::stripWhitespace)
    return(corpus)
}

cleaned_corpus <- cleanCorpus(corpus)
cat("Cleaned corpus size:", {length(cleaned_corpus)})

```

## Create DTM and explore data

```{r dtm}

dtm <- tm::DocumentTermMatrix(cleaned_corpus)
cat("DTM dimensions:", {dim(dtm)})

```

## Check for empty documents

```{r empty docs}

dtm_matrix <- as.matrix(dtm)
row_sums <- Matrix::rowSums(dtm_matrix)
cat("Document term counts:", {row_sums})

```

## Normalize the matrix

```{r normalize matrix}

dtm_normalized <- scale(dtm_matrix)

```

## Perform k-means Clustering with k=3
```{r k-means clustering}

set.seed(123)
kmeans_result <- stats::kmeans(dtm_normalized, centers = 3, nstart = 25)

```

## Create visualization using PCA - Principal Component Analysis

```{r PCA}

pca_result <- stats::prcomp(dtm_normalized)
cluster_plot_data <- data.frame(
    PC1 = pca_result$x[,1],
    PC2 = pca_result$x[,2],
    Cluster = factor(kmeans_result$cluster),
    Category = categories
)

# create scatter plot with enhanced visibility
ggplot2::ggplot(cluster_plot_data, ggplot2::aes(PC1, PC2, color = Category, shape = Cluster)) +
    ggplot2::geom_point(size = 5, alpha = 0.7) + 
    ggplot2::labs(title = "Document Clusters",
                    subtitle = "Sports, Technology, and Food Articles",
                    x = "First Principal Component",
                    y = "Second Principal Component") +
    ggthemes::theme_economist_white() +
    ggplot2::theme(legend.position = "right",
            plot.title = ggplot2::element_text(size = 14, face = "bold"),
            legend.text = ggplot2::element_text(size = 10))

```

## Print verification of cluster assignments

```{r cluster assignments}

print("Cluster assignments by category:")
print(table(Category = categories, Cluster = kmeans_result$cluster))

```

## Print top terms for each cluster

```{r top terms}

print("Top terms per cluster:")
terms <- colnames(dtm_matrix)
centers <- kmeans_result$centers
for(i in 1:3) {
    cat(paste0("\nCluster ", i, " top terms:\n"))
    top_indices <- order(centers[i,], decreasing = TRUE)[1:10]
    print(terms[top_indices])
}

```

# Part 2: Complex Text Clustering in R with a K-Means Algorithm

## Prepare the Data

```{r data prep 2}

austen_books <- janeaustenr::austen_books() %>%
    dplyr::group_by(book) %>%
    dplyr::mutate(
        # Add line numbers within each book
        linenumber = row_number(),
        # Create chapter groups
        chapter = cumsum(stringr::str_detect(text, stringr::regex("^chapter [\\dIVXLC]", ignore_case = TRUE)) | 
            stringr::str_detect(text, stringr::regex("^[\\dIVXLC]+", ignore_case = TRUE)))) %>%
    dplyr::ungroup()

```

## Organize by Chapter Words and Create Document-Term Matrix

```{r dtm 2}

stop_words <- tm::stopwords("english")

by_chapter_word <- austen_books %>%
    dplyr::filter(chapter > 0) %>%
    tidyr::unite(document, book, chapter) %>%
    tidytext::unnest_tokens(word, text)

# remove stop words and count words
word_counts <- by_chapter_word %>%
    dplyr::anti_join(tibble::tibble(word = stop_words), by = "word") %>% # changed anti-join
    dplyr::count(document, word, sort = TRUE) %>%
    dplyr::ungroup()

# create a document-term matrix
chapters_dtm <- word_counts %>%
    tidytext::cast_dtm(document, word, n)

```

## Let’s try to reduce the size of the dtm

```{r dtm dim 2}

dim(chapters_dtm)

# remove sparse terms before converting to matrix
chapters_dtm_reduced <- tm::removeSparseTerms(chapters_dtm, sparse = 0.99)
dim(chapters_dtm_reduced)

# now try the conversion with the reduced matrix
chapters_matrix <- as.matrix(chapters_dtm_reduced)
chapters_normalized <- scale(chapters_matrix)

```

## Determine Optimal Number of clusters with Elbow Method

```{r cluster optimization}

wss <- function(k) {
    stats::kmeans(chapters_normalized, k, nstart = 25)$tot.withinss
}

# Calculate WSS for k=1 to k=10
k.values <- 1:10
wss_values <- purrr::map_dbl(k.values, wss)


# Plot elbow curve
elbow_plot <- tibble::tibble(k = k.values, wss = wss_values) %>%
    ggplot2::ggplot(ggplot2::aes(k, wss)) +
    ggplot2::geom_line() +
    ggplot2::geom_point() +
    ggplot2::labs(title = "Elbow Method for Optimal k",
                    x = "Number of Clusters (k)",
                    y = "Total Within-cluster Sum of Squares") +
    ggthemes::theme_economist_white()

elbow_plot

```