---
title: "Lab 9"
author: "Conie OMalley"
date: "`r Sys.Date()`"
format: 
  pdf: 
    titlepage: true
    toc: true
    toc-depth: 2
    code-block-wrap: true
    number-sections: true
execute:
  echo: true
  freeze: false
  error: false
  jupyter: python3
python:
  version: /Users/coniecakes/anaconda3/envs/datascience/bin/python
---

# Pre Lab
```{r libraries}

required_packages <- c("tidyverse", "quanteda", "readtext", "stm", "stminsights", "wordcloud", "gsl", "topicmodels", 
                        "caret", "gutenbergr", "tidytext", "quanteda.textmodels", "tm", "igraph", "ggraph", "widyr", 
                        "jsonlite", "factoextra", "janeaustenr", "cluster", "SnowballC", "proxy", "stringr", "textclean",
                        "dendextend", "ggdendro", "plotly", "reticulate")

```

## Deliverable 1: Get your working directory and paste below:

"/Users/coniecakes/Library/CloudStorage/OneDrive-Personal/001. Documents - Main/023. Programming Tools/R Studio/AdvTxtAnalytics"

# Part 1: Introduction to K-Means Clustering with Synthetic Data

```{r}

texts <- c(
    # Sports articles
    "The football team scored three touchdowns. The quarterback threw perfect passes.",
    "The basketball team won with slam dunks and three-point shots.",
    "The baseball pitcher threw a perfect game with many strikeouts.",
    "The soccer team scored two goals and won the championship.",
    # Technology articles
    "The computer system processed data using advanced algorithms and databases.",
    "The network router managed bandwidth through secure encryption protocols.",
    "The software code executed functions through compiled programming syntax.",
    "The digital platform integrated APIs with cloud computing architecture.",
    # Food articles
    "The professional chef created sauces and prepared gourmet dishes.",
    "The master baker produced artisan breads and pastries daily.",
    "The culinary team seasoned meats and roasted vegetables.",
    "The kitchen staff garnished plates and plated entrees."
)

```

## Add category labels

```{r categories}

categories <- rep(c("Sports", "Technology", "Food"), each = 4)
cat("Initial document count:", {length(texts)})

```

# Create corpus and preprocess text

```{r corpus}

corpus <- tm::Corpus(tm::VectorSource(texts))
cat("Corpus size:", {length(corpus)})

cleanCorpus <- function(corpus) {
    corpus <- tm::tm_map(corpus, tm::content_transformer(tolower))
    corpus <- tm::tm_map(corpus, tm::removePunctuation)
    corpus <- tm::tm_map(corpus, tm::removeNumbers)
    corpus <- tm::tm_map(corpus, tm::removeWords, tm::stopwords("english"))
    corpus <- tm::tm_map(corpus, tm::stripWhitespace)
    return(corpus)
}

cleaned_corpus <- cleanCorpus(corpus)
cat("Cleaned corpus size:", {length(cleaned_corpus)})

```

## Create DTM and explore data

```{r dtm}

dtm <- tm::DocumentTermMatrix(cleaned_corpus)
cat("DTM dimensions:", {dim(dtm)})

```

## Check for empty documents

```{r empty docs}

dtm_matrix <- as.matrix(dtm)
row_sums <- Matrix::rowSums(dtm_matrix)
cat("Document term counts:", {row_sums})

```

## Normalize the matrix

```{r normalize matrix}

dtm_normalized <- scale(dtm_matrix)

```

## Perform k-means Clustering with k=3
```{r k-means clustering}

set.seed(123)
kmeans_result <- stats::kmeans(dtm_normalized, centers = 3, nstart = 25)

```

## Create visualization using PCA - Principal Component Analysis

```{r PCA}

pca_result <- stats::prcomp(dtm_normalized)
cluster_plot_data <- data.frame(
    PC1 = pca_result$x[,1],
    PC2 = pca_result$x[,2],
    Cluster = factor(kmeans_result$cluster),
    Category = categories
)

# create scatter plot with enhanced visibility
ggplot2::ggplot(cluster_plot_data, ggplot2::aes(PC1, PC2, color = Category, shape = Cluster)) +
    ggplot2::geom_point(size = 5, alpha = 0.7) + 
    ggplot2::labs(title = "Document Clusters",
                    subtitle = "Sports, Technology, and Food Articles",
                    x = "First Principal Component",
                    y = "Second Principal Component") +
    ggthemes::theme_economist_white() +
    ggplot2::theme(legend.position = "right",
            plot.title = ggplot2::element_text(size = 14, face = "bold"),
            legend.text = ggplot2::element_text(size = 10))

```

## Print verification of cluster assignments

```{r cluster assignments}

print("Cluster assignments by category:")
print(table(Category = categories, Cluster = kmeans_result$cluster))

```

## Print top terms for each cluster

```{r top terms}

print("Top terms per cluster:")
terms <- colnames(dtm_matrix)
centers <- kmeans_result$centers
for(i in 1:3) {
    cat(paste0("\nCluster ", i, " top terms:\n"))
    top_indices <- order(centers[i,], decreasing = TRUE)[1:10]
    print(terms[top_indices])
}

```

# Part 2: Complex Text Clustering in R with a K-Means Algorithm

## Prepare the Data

```{r data prep 2}

austen_books <- janeaustenr::austen_books() %>%
    dplyr::group_by(book) %>%
    dplyr::mutate(
        # Add line numbers within each book
        linenumber = row_number(),
        # Create chapter groups
        chapter = cumsum(stringr::str_detect(text, stringr::regex("^chapter [\\dIVXLC]", ignore_case = TRUE)) | 
            stringr::str_detect(text, stringr::regex("^[\\dIVXLC]+", ignore_case = TRUE)))) %>%
    dplyr::ungroup()

```

## Organize by Chapter Words and Create Document-Term Matrix

```{r dtm 2}

stop_words <- tm::stopwords("english")

by_chapter_word <- austen_books %>%
    dplyr::filter(chapter > 0) %>%
    tidyr::unite(document, book, chapter) %>%
    tidytext::unnest_tokens(word, text)

# remove stop words and count words
word_counts <- by_chapter_word %>%
    dplyr::anti_join(tibble::tibble(word = stop_words), by = "word") %>% # changed anti-join
    dplyr::count(document, word, sort = TRUE) %>%
    dplyr::ungroup()

# create a document-term matrix
chapters_dtm <- word_counts %>%
    tidytext::cast_dtm(document, word, n)

```

## Let’s try to reduce the size of the dtm

```{r dtm dim 2}

dim(chapters_dtm)

# remove sparse terms before converting to matrix
chapters_dtm_reduced <- tm::removeSparseTerms(chapters_dtm, sparse = 0.99)
dim(chapters_dtm_reduced)

# now try the conversion with the reduced matrix
chapters_matrix <- as.matrix(chapters_dtm_reduced)
chapters_normalized <- scale(chapters_matrix)

```

## Determine Optimal Number of clusters with Elbow Method

```{r cluster optimization}

wss <- function(k) {
    stats::kmeans(chapters_normalized, k, nstart = 25)$tot.withinss
}

# Calculate WSS for k=1 to k=10
k.values <- 1:10
wss_values <- purrr::map_dbl(k.values, wss)


# Plot elbow curve
elbow_plot <- tibble::tibble(k = k.values, wss = wss_values) %>%
    ggplot2::ggplot(ggplot2::aes(k, wss)) +
    ggplot2::geom_line() +
    ggplot2::geom_point() +
    ggplot2::labs(title = "Elbow Method for Optimal k",
                    x = "Number of Clusters (k)",
                    y = "Total Within-cluster Sum of Squares") +
    ggthemes::theme_economist_white()

elbow_plot

```

## Perform k-means clustering with k=3

```{r k means clustering}

# k=3 means clustering
set.seed(123) # for reproducibility
kmeans_3 <- stats::kmeans(chapters_normalized, centers = 3, nstart = 25)

# Perform k-means clustering with k=6
set.seed(123)
kmeans_6 <- stats::kmeans(chapters_normalized, centers = 6, nstart = 25)

```

## Create visualization and analysis for both

```{r k means visualization}

# First, let's look at which chapters are assigned to which clusters
cluster_assignments_3 <- word_counts %>%
    tidyr::separate(document, c("book", "chapter"), sep = "_") %>%
    dplyr::distinct(book, chapter) %>%
    dplyr::mutate(cluster = kmeans_3$cluster[as.numeric(factor(paste0(book, chapter, sep = "_")))])

cluster_assignments_6 <- word_counts %>%
    tidyr::separate(document, c("book", "chapter"), sep = "_") %>%
    dplyr::distinct(book, chapter) %>%
    dplyr::mutate(cluster = kmeans_6$cluster[as.numeric(factor(paste(book, chapter, sep = "_")))])

```

## Create heatmaps for both k=3 and k=6
```{r k means heatmaps}

heatmap_3 <- cluster_assignments_3 %>%
    dplyr::count(book, cluster) %>%
    ggplot2::ggplot(ggplot2::aes(factor(cluster), book, fill = n)) +
        ggplot2::geom_tile() +
        ggplot2::scale_fill_gradient(low = "white", high = "steelblue") +
        ggplot2::labs(title = "Cluster Distribution (k=3)",
                        x = "Cluster",
                        y = "Book",
                        fill = "Number of Chapters") +
        ggthemes::theme_economist_white()

heatmap_6 <- cluster_assignments_6 %>%
    dplyr::count(book, cluster) %>%
    ggplot2::ggplot(ggplot2::aes(factor(cluster), book, fill = n)) +
        ggplot2::geom_tile() +
        ggplot2::scale_fill_gradient(low = "white", high = "steelblue") +
        ggplot2::labs(title = "Cluster Distribution (k=6)",
                        x = "Cluster",
                        y = "Book",
                        fill = "Number of Chapters") +
        ggthemes::theme_economist_white()

```

## Print Summary Statistics
```{r summary statistics}

print("Cluster sizes for k=3:")
print(table(cluster_assignments_3$cluster))
print("Cluster composition for k=3:")
print(table(cluster_assignments_3$book, cluster_assignments_3$cluster))
print("Cluster sizes for k=6:")
print(table(cluster_assignments_6$cluster))
print("Cluster composition for k=6:")
print(table(cluster_assignments_6$book, cluster_assignments_6$cluster))

```

## Display Heatmaps for both K Values
```{r heatmaps}

print(heatmap_3)
print(heatmap_6)

```

## Calculate and Display Characteristic Words in Each Cluster (k=3)
```{r characteristics k3}

cluster_words_3 <- word_counts %>%
    tidyr::separate(document, c("book", "chapter"), sep = "_") %>%
    dplyr::mutate(cluster = kmeans_3$cluster[as.numeric(factor(paste0(book, chapter, sep = "_")))]) %>%
    dplyr::group_by(cluster, word) %>%
    dplyr::summarise(total = sum(n), .groups = 'drop') %>%
    dplyr::group_by(cluster) %>%
    dplyr::slice_max(total, n = 10) %>% # Get top 10 words for each cluster
    dplyr::arrange(cluster, desc(total))

print("Top words in each cluster (k=3):")
print(cluster_words_3)

```

## Calculate and Display Characteristic Words in Each Cluster (k=6)
```{r characteristic-words-k6}

cluster_words_6 <- word_counts %>%
    tidyr::separate(document, c("book", "chapter"), sep = "_") %>%
    dplyr::mutate(cluster = kmeans_6$cluster[as.numeric(factor(paste(book, chapter, sep = "_")))]) %>%
    dplyr::group_by(cluster, word) %>%
    dplyr::summarise(total = sum(n), .groups = 'drop') %>%
    dplyr::group_by(cluster) %>%
    dplyr::slice_max(total, n = 15) %>% # Get top 15 words for each cluster
    dplyr::arrange(cluster, desc(total))

print("Top words in each cluster (k=6):")
print(cluster_words_6)

```

## Print the k=3 results in a more readable format
```{r}

cluster_words_3 %>%
    dplyr::group_by(cluster) %>%
    dplyr::summarise(words = paste0(word, collapse = ", ")) %>%
    dplyr::mutate(cluster = paste0("Cluster", cluster)) %>%
    print(width = Inf) # Print full width

```

# Part 3: Hierarchical (Tree) Clustering in R

## Load the Customer Complaint’s Dataset and Preprocess
```{r}

complaints <- utils::read.csv("/Users/coniecakes/Library/CloudStorage/OneDrive-Personal/001. Documents - Main/004_School/001. MBA_MS/008. Spring 2025/004. Advanced Text Analytics_ML/complaints.csv", stringsAsFactors = FALSE)
# Text preprocessing function
preprocess_text <- function(text) {
    text %>%
        # Convert to lowercase
        stringr::str_to_lower() %>%
        # Remove special characters and digits
        stringr::str_replace_all("[^[:alpha:][:space:]]", "") %>%
        # Remove extra whitespace
        stringr::str_trim() %>%
        stringr::str_squish()
}
# Apply preprocessing
complaints <- complaints %>%
    dplyr::mutate(processed_narrative = purrr::map_chr(consumer_complaint_narrative,
                                                        ~preprocess_text(as.character(.))))
    
# Remove missing values
complaints <- complaints %>%
    dplyr::filter(!is.na(processed_narrative))

# Create corpus
corpus <- tm::Corpus(tm::VectorSource(complaints$processed_narrative))

```

## Perform Hierarchical Clustering
```{r hierarchical clustering}

# Create document-term matrix
dtm_hierarchy <- tm::DocumentTermMatrix(corpus,
                                        control = list(
                                        weighting = tm::weightTfIdf,
                                        stopwords = TRUE,
                                        removeNumbers = TRUE,
                                        removePunctuation = TRUE,
                                        stemming = TRUE))

# Convert to matrix and remove sparse terms
dtm_matrix_hierarchy <- tm::removeSparseTerms(dtm_hierarchy, sparse = 0.99)
dtm_matrix_hierarchy <- as.matrix(dtm_matrix_hierarchy)

```

## Calculate Distance Matrix and Perform Hierarchical Clustering
```{r dist matrix and clustering}

# Using first 100 documents for visualization purposes
dist_matrix <- stats::dist(dtm_matrix_hierarchy[1:100,], method = "euclidean")

# Perform hierarchical clustering
hc <- stats::hclust(dist_matrix, method = "ward.D2")

```

## Visualize the Dendrogram
```{r dendrogram}

# Basic dendrogram
graphics::plot(hc, main = "Hierarchical Clustering Dendrogram",
                xlab = "Documents", ylab = "Height",
                cex = 0.6, hang = -1)

```

```{r enhance dendrogram}

# Enhanced dendrogram using dendextend
dend <- stats::as.dendrogram(hc)
dend <- dendextend::color_branches(dend, k = 5) # Color branches for 5 clusters

# Plot enhanced dendrogram
graphics::par(mar = c(5,5,3,1))
graphics::plot(dend,
                main = "Colored Dendrogram with 5 Clusters",
                ylab = "Height",
                leaflab = "none")

```

## Cut the Tree into Clusters and Analyze
```{r tree cluster analysis}

# Cut tree into 5 clusters
clusters <- dendextend::cutree(hc, k = 5)
# Add cluster assignments to the original data
complaints_subset <- complaints[1:100,] %>%
    dplyr::mutate(cluster = as.factor(clusters))
# Analyze clusters
cluster_summary <- complaints_subset %>%
    dplyr::group_by(cluster) %>%
    dplyr::summarise(
                n = n(),
                sample_complaints = list(utils::head(processed_narrative, 3))
)

# Print cluster summaries
for(i in 1:nrow(cluster_summary)) {
    cat("\nCluster", i, "\n")
    cat("Number of complaints:", cluster_summary$n[i], "\n")
    cat("Sample complaints:\n")
    print(unlist(cluster_summary$sample_complaints[i]))
    cat("\n")
}

```

```{r silhouette plot}

sil <- cluster::silhouette(clusters, dist_matrix)
graphics::plot(sil, main = "Silhouette Plot")

# Average silhouette width
avg_sil <- mean(sil[,3])
cat("Average silhouette width:", avg_sil)

```

```{r color silhouette plot}

# Plot silhouette analysis using factoextra
factoextra::fviz_silhouette(sil)

```

## Get Top Words for Each Cluster
```{r top words}

# Function to get top words for each cluster
get_top_words <- function(cluster_num, n_words = 10) {
    # Get documents in cluster
    cluster_docs <- complaints_subset %>%
        dplyr::filter(cluster == cluster_num) %>%
        dplyr::pull(processed_narrative)
        # Create DTM for cluster
        cluster_corpus <- tm::Corpus(tm::VectorSource(cluster_docs))
        cluster_dtm <- tm::DocumentTermMatrix(cluster_corpus,
                                                control = list(
                                                        weighting = tm::weightTfIdf,
                                                        stopwords = TRUE)
                                                        )
    # Get term frequencies
    term_freq <- colSums(as.matrix(cluster_dtm))
    top_terms <- sort(term_freq, decreasing = TRUE)[1:n_words]
    return(names(top_terms))
}

# Print top words for each cluster
for(i in 1:5) {
    cat("\nCluster", i, "top words:\n")
    print(get_top_words(i))
}

```

## Create Cluster Visualization Using Factoextra
```{r cluster viz}

# Perform PCA on the distance matrix for visualization
pca <- stats::prcomp(dtm_matrix_hierarchy[1:100,])

# Plot clusters in first two principal components
factoextra::fviz_cluster(list(data = pca$x[,1:2], cluster = clusters),
                                main = "Cluster Plot",
                                ellipse.type = "convex",
                                repel = TRUE,
                                show.clust.cent = TRUE)

```

## Compare Different Distance Metrics
```{r metrics comparison}

# Compare different distance metrics
distance_methods <- c("euclidean", "manhattan", "cosine")
dendlist <- list()
for(method in distance_methods) {
    dist_mat <- proxy::dist(dtm_matrix_hierarchy[1:100,], method = method)
    hc_temp <- stats::hclust(dist_mat, method = "ward.D2")
    dendlist[[method]] <- stats::as.dendrogram(hc_temp)
}
# Compare dendrograms
dend_list <- dendextend::dendlist(dendlist[[1]], dendlist[[2]], dendlist[[3]])
names(dend_list) <- distance_methods
dendextend::tanglegram(dend_list[[1]], dend_list[[2]])

```

## Convert Dendrogram to Plotly for Interactive Visualization
```{r plotly visual}

# Convert dendrogram to plotly
dend_data <- ggdendro::dendro_data(dend)

p <- ggplot2::ggplot(ggdendro::segment(dend_data)) +
                        ggplot2::geom_segment(ggplot2::aes(x = x, y = y, xend = xend, yend = yend)) +
                        ggplot2::labs(title = "Dendrogram Visualization") +
                        ggthemes::theme_economist_white()
                        

plotly::ggplotly(p)

```

# Part 4: Introduction to Topic Modeling

## Get AP Data and Build a DTM to submit for the LDA model

```{r ap data load}

data("AssociatedPress")
AssociatedPress

ap_lda <- topicmodels::LDA(AssociatedPress, k=8, control = list(seed=1234))
ap_lda

```

## Explore and Interpret the Model

```{r lda model interpretation}

# Step 1: Extract beta matrix from the LDA model
beta_matrix <- exp(ap_lda@beta)  # Convert log probabilities to probabilities

# Step 2: Retrieve the terms (vocabulary) from the DocumentTermMatrix
terms <- colnames(AssociatedPress)

# Step 3: Validate the beta matrix dimensions with vocabulary size
if (ncol(beta_matrix) != length(terms)) {
  stop("Mismatch between number of columns in beta_matrix and the length of terms.")
}

# Step 4: Convert beta matrix to data frame and assign topic identifiers
beta_df <- as.data.frame(beta_matrix)
beta_df$topic <- 1:nrow(beta_df)

# Step 5: Pivot the data to a long format correctly
ap_topics_long <- beta_df %>%
  tidyr::pivot_longer(
    cols = -topic,      # Keep the topic column static
    names_to = "term_index",  # Create index for terms
    values_to = "beta"
  ) %>%
  dplyr::mutate(
    term_index = as.integer(gsub("V", "", term_index)),  # Extract integer index
    term = terms[term_index]  # Map term names using the index
  ) %>%
  dplyr::select(topic, term, beta)

# Step 6: Display the tidy data frame
print(head(ap_topics_long, 20))

# Step 7: Extract the top terms for each topic using dplyr
ap_top_terms <- ap_topics_long %>%
  dplyr::group_by(topic) %>%
  dplyr::top_n(10, beta) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(topic, dplyr::desc(beta))

# Display the top terms for each topic
print(head(ap_top_terms, 20))

```

## Visualize the Top Terms

```{r top terms visualization}

ap_top_terms %>%
    dplyr::mutate(term = reorder(term, beta)) %>%
    ggplot2::ggplot(ggplot2::aes(term, beta, fill = factor(topic))) +
    ggplot2::geom_col(show.legend = FALSE) +
    ggplot2::facet_wrap(~ topic, scales = "free") +
    ggplot2::coord_flip() + 
    ggthemes::theme_economist_white() +
    ggplot2::labs(title = "Top Terms by Beta and Topic",
                 x = "Term",)

```

```{r beta spread}

beta_spread <- ap_topics_long %>%
  dplyr::mutate(topic = paste0("topic", topic)) %>%  # Create a topic identifier
  tidyr::pivot_wider(
    names_from = topic,        # Pivot topic names to separate columns
    values_from = beta         # Populate those columns with beta values
  ) %>%
  dplyr::filter(!is.na(topic1) & !is.na(topic2) & (topic1 > 0.001 | topic2 > 0.001)) %>%
  dplyr::mutate(log_ratio = log2(topic2 / topic1))

beta_spread

```

```{r gamma matrix}

gamma_matrix <- exp(ap_lda@gamma)  # Convert log probabilities to probabilities

document_count <- nrow(AssociatedPress)

documents <- 1:document_count  # This is a placeholder, replace with actual document IDs if available

gamma_df <- as.data.frame(gamma_matrix)
gamma_df$document <- documents

ap_documents_long <- gamma_df %>%
  tidyr::pivot_longer(
    cols = -document,      # Keep the document column static
    names_to = "topic_index",  # Create index for topics
    values_to = "gamma"
  ) %>%
  dplyr::mutate(
    topic_index = as.integer(gsub("V", "", topic_index)),  # Extract integer index
    topic = topic_index  # Direct mapping for simplicity
  ) %>%
  dplyr::select(document, topic, gamma)

print(head(ap_documents_long, 20))

```


# Part 5: Topic Modelling Sanity Check with Project Gutenberg

## Collect and Prepare the Data

```{r gutenberg books download}

options(gutenberg_mirror = "https://gutenberg.pglaf.org")

books <- gutenbergr::gutenberg_download(c(1259, # Great Expectations (Dickens)
                                            36, # The War of the Worlds (Wells)
                                            98, # A Tale of Two Cities (Dickens)
                                            74)) # The Adventures of Tom Sawyer (Twain)


# Create labels for the books
books <- books %>%
    dplyr::mutate(title = dplyr::case_when(
                                    gutenberg_id == 1259 ~ "Great Expectations",
                                    gutenberg_id == 36 ~ "The War of the Worlds",
                                    gutenberg_id == 98 ~ "Tale of Two Cities",
                                    gutenberg_id == 74 ~ "Tom Sawyer",
                                    TRUE ~ NA_character_
))

books

```

```{r chapters check}

by_chapter <- books %>%
    dplyr::group_by(title) %>%
    dplyr::mutate(
# Enhanced chapter detection for different formats
            chapter = cumsum(
                        stringr::str_detect(text, stringr::regex("^chapter [\\dIVXLC]", ignore_case = TRUE)) |
                        stringr::str_detect(text, stringr::regex("^[\\dIVXLC]+\\.", ignore_case = TRUE))
        )
    ) %>%
    dplyr::ungroup() %>%
    dplyr::filter(chapter > 0) %>%
    # Remove Project Gutenberg headers/footers
    dplyr::filter(!stringr::str_detect(text, "Project Gutenberg")) %>%
    dplyr::filter(!stringr::str_detect(text, "\\*\\*\\* START OF")) %>%
    dplyr::filter(!stringr::str_detect(text, "\\*\\*\\* END OF"))

chapter_counts <- by_chapter %>%
    dplyr::group_by(title) %>%
    dplyr::summarize(chapters = dplyr::n_distinct(chapter))

print("Chapters per book:")
print(chapter_counts)

```

```{r tokenize gutenberg books}

# Tokenize and remove stop words
by_chapter_word <- by_chapter %>%
    tidyr::unite(document, title, chapter) %>%
    tidytext::unnest_tokens(word, text)
# Assuming 'word' is the column in 'by_chapter_word' representing the individual words
stop_words_df <- tibble(word = stop_words)

word_counts <- by_chapter_word %>%
  dplyr::anti_join(stop_words_df, by = "word") %>%  # Use 'by = "word"' to specify the join column
  dplyr::count(document, word, sort = TRUE) %>%
  dplyr::ungroup()

```

## Prepare the Data: Convert the Tibble to a DTM
```{r gutenberg dtm}

# Create document-term matrix
chapters_dtm <- word_counts %>%
    tidytext::cast_dtm(document, word, n)

chapters_dtm

```

## Create the LDA Topic Model
```{r gutenberg lda model}

# Fit LDA model
chapters_lda <- topicmodels::LDA(chapters_dtm, k = 4, control = list(seed = 1234))
chapters_lda

```

```{r gutenberg beta matrix}

chapters_beta_matrix <- exp(chapters_lda@beta)  # Convert log probabilities to probabilities

chapters_terms <- colnames(chapters_dtm)

chapters_beta_df <- as.data.frame(chapters_beta_matrix)
chapters_beta_df$topic <- 1:nrow(chapters_beta_df)

chapters_topics_long <- chapters_beta_df %>%
  tidyr::pivot_longer(
    cols = -topic,      # Keep the topic column static
    names_to = "term_index",  # Create index for terms
    values_to = "beta"
  ) %>%
  dplyr::mutate(
    term_index = as.integer(gsub("V", "", term_index)),  # Extract integer index
    term = chapters_terms[term_index]  # Map term names using the index
  ) %>%
  dplyr::select(topic, term, beta)

print(head(chapters_topics_long, 20))

chapters_top_terms <- chapters_topics_long %>%
  dplyr::group_by(topic) %>%
  dplyr::top_n(10, beta) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(topic, dplyr::desc(beta))

# Display the top terms for each topic
print(head(chapters_top_terms, 20))

```

```{r gutenberg dtm visualization}

# Create visualization of top terms
top_terms_plot <- chapters_top_terms %>%
    dplyr::mutate(term = tidytext::reorder_within(term, beta, topic)) %>%
    ggplot2::ggplot(ggplot2::aes(beta, term, fill = factor(topic))) +
    ggplot2::geom_col(show.legend = FALSE) +
    ggplot2::facet_wrap(~topic, scales = "free_y") +
    tidytext::scale_y_reordered() +
    ggplot2::labs(title = "Top 10 Terms in Each Topic",
                    x = "Beta Score",
                    y = NULL) +
    ggthemes::theme_economist_white()

top_terms_plot

```

```{r gutenberg gamma matrix}

gamma_matrix <- exp(chapters_lda@gamma)  # Convert log probabilities to probabilities

document_ids <- rownames(chapters_dtm)  # Placeholder, adjust to your source

gamma_df <- as.data.frame(gamma_matrix)
gamma_df$document <- document_ids

chapters_gamma <- gamma_df %>%
    tidyr::pivot_longer(
        cols = -document,          # All columns except 'document'
        names_to = "topic_index",  # Create an index for topics
        values_to = "gamma"        # Store probabilities in 'gamma'
    ) %>%
    dplyr::mutate(
        topic_index = as.integer(gsub("V", "", topic_index)),
        topic = topic_index
    )
# Separate document field back into title and chapter
chapters_gamma <- chapters_gamma %>%
    tidyr::separate(document, c("title", "chapter"), sep = "_", convert = TRUE)
# Calculate average topic probabilities per book
book_topics <- chapters_gamma %>%
    dplyr::group_by(title, topic) %>%
    dplyr::summarise(gamma = mean(gamma)) %>%
    dplyr::ungroup()

```

## Create a Per-Document Classification
```{r gutenberg document classification}

# Create book-topic heatmap
book_topics_plot <- book_topics %>%
    ggplot2::ggplot(ggplot2::aes(factor(topic), title, fill = gamma)) +
    ggplot2::geom_tile() +
    ggplot2::scale_fill_gradient(low = "white", high = "red") +
    ggplot2::labs(title = "Topic Distribution Across Books",
                    x = "Topic",
                    y = "Book",
                    fill = "Gamma") +
    ggthemes::theme_economist_white()

book_topics_plot


```

```{r gutenberg summaries}

# Print top terms for each topic
print("Top terms for each topic:")
chapters_top_terms %>%
    dplyr::group_by(topic) %>%
    dplyr::slice_head(n = 10) %>%
    dplyr::arrange(topic, -beta) %>%

print(n = 40)

# Print topic distribution by book
print("Topic distribution by book:")
book_topics %>%
    tidyr::pivot_wider(names_from = topic,
                        values_from = gamma,
                        names_prefix = "Topic ") %>%
    dplyr::arrange(desc(`Topic 1`)) %>%
    print()

# Calculate and print the most distinctive words for each book
distinctive_words <- by_chapter_word %>%
    tidyr::separate(document, c("title", "chapter"), sep = "_") %>%
    dplyr::anti_join(stop_words_df) %>%
    dplyr::count(title, word) %>%
    tidytext::bind_tf_idf(word, title, n) %>%
    dplyr::group_by(title) %>%
    dplyr::slice_max(tf_idf, n = 10) %>%
    dplyr::ungroup()

print("Most distinctive words for each book:")
distinctive_words %>%
    dplyr::arrange(title, desc(tf_idf)) %>%
    print(n = 40)

```

# Part 6: Clustering in Python: K-Means, DBSCAN, and Hierarchical

## Importing Libraries and Loading Data
```{python imports}

import numpy as np
import pandas as pd
import matplotlib.pyplot as plt
import seaborn as sns
from sklearn.datasets import fetch_20newsgroups
from sklearn.feature_extraction.text import TfidfVectorizer
from sklearn.cluster import KMeans, DBSCAN, AgglomerativeClustering
from sklearn.decomposition import TruncatedSVD
from sklearn.metrics import silhouette_score, calinski_harabasz_score
from sklearn.preprocessing import normalize
from scipy.cluster.hierarchy import dendrogram, linkage
from wordcloud import WordCloud
import umap
import warnings
warnings.filterwarnings('ignore')

```

```{python data load}

categories = ['comp.graphics', 'rec.sport.baseball', 'sci.space', 'talk.politics.guns']
newsgroups = fetch_20newsgroups(categories=categories, remove=('headers', 'footers', 'quotes'))

```

## Creating TF-IDF Matrix
```{python py tfidf matrix}

# Create TF-IDF matrix
vectorizer = TfidfVectorizer(max_features=5000,
stop_words='english',
max_df=0.95,
min_df=2)
tfidf_matrix = vectorizer.fit_transform(newsgroups.data)

```

## Dimensionality Reduction for Visualization
```{python dim reduction viz}

# Using TruncatedSVD (LSA) first
n_components = 100
lsa = TruncatedSVD(n_components=n_components)
doc_vectors_lsa = lsa.fit_transform(tfidf_matrix)
# Further reduce to 2D using UMAP for visualization
reducer = umap.UMAP(random_state=42)
doc_vectors_2d = reducer.fit_transform(doc_vectors_lsa)

```

## Preparing for and Conducting K-means Clustering
```{python py k means clustering}

def perform_kmeans_clustering(vectors, n_clusters=4):
    kmeans = KMeans(n_clusters=n_clusters, random_state=42)
    cluster_labels = kmeans.fit_predict(vectors)
    # Get cluster centers
    if vectors.shape[1] > 2:
        centers_2d = reducer.transform(kmeans.cluster_centers_)
    else:
        centers_2d = kmeans.cluster_centers_
    return kmeans, cluster_labels, centers_2d

# Perform k-means clustering
kmeans_model, kmeans_labels, centers = perform_kmeans_clustering(doc_vectors_lsa)

```

## Visualize K-means Clusters
```{python py k means cluster viz}

plt.figure(figsize=(10, 8))
scatter = plt.scatter(doc_vectors_2d[:, 0], doc_vectors_2d[:, 1],
c=kmeans_labels, cmap='viridis', alpha=0.6)
plt.scatter(centers[:, 0], centers[:, 1], c='red', marker='x', s=200, linewidths=3)
plt.title('K-means Clustering of News Articles')
plt.colorbar(scatter)
plt.show()

```

## Hierarchical Clustering in Python
```{python py hierarchical clustering}

def perform_hierarchical_clustering(vectors, n_clusters=4):
    # Perform hierarchical clustering
    hc = AgglomerativeClustering(n_clusters=n_clusters)
    hc_labels = hc.fit_predict(vectors)
    # Create linkage matrix for dendrogram
    linkage_matrix = linkage(vectors, method='ward')
    return hc, hc_labels, linkage_matrix

# Perform hierarchical clustering
hc_model, hc_labels, linkage_matrix = perform_hierarchical_clustering(doc_vectors_lsa)

```

## Plot Dendrogram
```{python py dendrogram}

plt.figure(figsize=(12, 8))
dendrogram(linkage_matrix)
plt.title('Hierarchical Clustering Dendrogram')
plt.xlabel('Sample Index')
plt.ylabel('Distance')
plt.show()

```

## DBSCAN Clustering
```{python py dbscan clustering}

def perform_dbscan_clustering(vectors, eps=0.5, min_samples=5):
    dbscan = DBSCAN(eps=eps, min_samples=min_samples)
    dbscan_labels = dbscan.fit_predict(vectors)
    return dbscan, dbscan_labels

# Perform DBSCAN clustering
dbscan_model, dbscan_labels = perform_dbscan_clustering(doc_vectors_lsa)

```

## Cluster Analysis and Evaluation
```{python py cluster analysis}

def analyze_clusters(vectors, labels, vectorizer):
    # Convert sparse matrix to dense array
    vectors_dense = vectors.toarray() if hasattr(vectors, 'todense') else vectors

    # Calculate silhouette score
    sil_score = silhouette_score(vectors_dense, labels)
    # Calculate Calinski-Harabasz score
    ch_score = calinski_harabasz_score(vectors_dense, labels)
    
    # Get top terms per cluster
    n_terms = 10
    cluster_terms = {}
    # Original feature names
    feature_names = np.array(vectorizer.get_feature_names_out())
    # For each cluster
    for cluster_id in np.unique(labels):
        if cluster_id != -1:  # Ignore noise points from DBSCAN
            # Get points in cluster
            cluster_docs = vectors[labels == cluster_id].toarray() if hasattr(vectors, 'todense') else vectors[labels == cluster_id]
            # Calculate mean tfidf scores for each term
            centroid = cluster_docs.mean(axis=0)
            # Get top terms
            top_term_indices = centroid.argsort()[-n_terms:][::-1]
            top_terms = feature_names[top_term_indices]
            cluster_terms[f'Cluster {cluster_id}'] = top_terms

    return {
        'silhouette_score': sil_score,
        'calinski_harabasz_score': ch_score,
        'top_terms': cluster_terms
    }

# Example usage
kmeans_analysis = analyze_clusters(tfidf_matrix, kmeans_labels, vectorizer)
hc_analysis = analyze_clusters(tfidf_matrix, hc_labels, vectorizer)
dbscan_analysis = analyze_clusters(tfidf_matrix, dbscan_labels, vectorizer)

```

```{python py cluster analysis 2}

# Elbow Method for K-means
def plot_elbow_curve(vectors, max_k=10):
    distortions = []
    K = range(1, max_k+1)
    for k in K:
        kmeans = KMeans(n_clusters=k, random_state=42)
        kmeans.fit(vectors)
        distortions.append(kmeans.inertia_)

    plt.figure(figsize=(10, 6))
    plt.plot(K, distortions, 'bx-')
    plt.xlabel('k')
    plt.ylabel('Distortion')
    plt.title('The Elbow Method for Optimal k')
    plt.show()

plot_elbow_curve(doc_vectors_lsa)

```

```{python py cluster analysis 3}

# Generate Word Clouds for Clusters

def generate_cluster_wordcloud(vectors, labels, vectorizer, cluster_id):
    # Get documents in cluster
    cluster_docs = vectors[labels == cluster_id]
    # Get feature names
    feature_names = vectorizer.get_feature_names_out()
    # Calculate mean tfidf scores
    centroid = cluster_docs.mean(axis=0).A1 if hasattr(cluster_docs, 'todense') else cluster_docs.mean(axis=0)
    # Create word frequency dictionary
    word_freq = {feature_names[i]: centroid[i]
                 for i in range(len(feature_names))
                 if centroid[i] > 0}

    # Generate word cloud
    wordcloud = WordCloud(width=800, height=400,
                          background_color='white').generate_from_frequencies(word_freq)

    plt.figure(figsize=(10, 5))
    plt.imshow(wordcloud, interpolation='bilinear')
    plt.axis('off')
    plt.title(f'Word Cloud for Cluster {cluster_id}')
    plt.show()

# Generate word clouds for K-means clusters
for cluster_id in np.unique(kmeans_labels):
    generate_cluster_wordcloud(tfidf_matrix, kmeans_labels, vectorizer, cluster_id)

```

```{python py cluster analysis 4}

# Silhouette Analysis Visualization
def plot_silhouette_analysis(vectors, labels):
    plt.figure(figsize=(10, 6))
    silhouette_avg = silhouette_score(vectors, labels)
    sample_silhouette_values = silhouette_samples(vectors, labels)
    y_lower = 10
    for i in range(len(np.unique(labels))):
        ith_cluster_silhouette_values = sample_silhouette_values[labels == i]
        ith_cluster_silhouette_values.sort()

        size_cluster_i = ith_cluster_silhouette_values.shape[0]
        y_upper = y_lower + size_cluster_i

        plt.fill_betweenx(np.arange(y_lower, y_upper),
                            0, ith_cluster_silhouette_values,
                            alpha=0.7)
        y_lower = y_upper + 10

    plt.title("Silhouette Analysis")
    plt.xlabel("Silhouette Coefficient")
    plt.ylabel("Cluster")
    plt.axvline(x=silhouette_avg, color="red", linestyle="--")
    plt.show()

# Plot silhouette analysis for K-means
from sklearn.metrics import silhouette_samples

plot_silhouette_analysis(doc_vectors_lsa, kmeans_labels)

# Compare with Topic Modeling
from sklearn.decomposition import LatentDirichletAllocation

def compare_clustering_with_lda(vectors, cluster_labels, n_topics=4):
    # Perform LDA
    lda = LatentDirichletAllocation(n_components=n_topics, random_state=42)
    doc_topics = lda.fit_transform(vectors)
    topic_labels = doc_topics.argmax(axis=1)

    # Create confusion matrix
    confusion_matrix = pd.crosstab(cluster_labels, topic_labels,
                                    normalize='index')

    # Plot heatmap
    plt.figure(figsize=(10, 8))
    sns.heatmap(confusion_matrix, annot=True, cmap='YlOrRd')
    plt.title('Cluster vs Topic Distribution')
    plt.xlabel('Topic')
    plt.ylabel('Cluster')
    plt.show()

# Compare K-means clusters with LDA topics
compare_clustering_with_lda(tfidf_matrix, kmeans_labels)

```

# Part 7: LDA Topic Modelling in Python

## Importing Libraries and Loading Data
```{python libraries 2}

# Import necessary libraries
import nltk
from gensim import corpora, models
import matplotlib.pyplot as plt
# Download NLTK data (run this only once)
nltk.download('reuters')

```

```{python data import 2}

# Load the Reuters dataset from NLTK
from nltk.corpus import reuters

# Select a subset of documents for simplicity (e.g., first 100 documents)
docs = [reuters.raw(fileid) for fileid in reuters.fileids()[:100]]

```

## Preprocessing the Data
```{python data preprocessing 2}

from nltk.tokenize import word_tokenize
from nltk.corpus import stopwords
import string

# Function to preprocess text
def preprocess(text):
    # Convert text to lowercase, tokenize, remove punctuation and stopwords
    tokens = word_tokenize(text.lower())
    tokens = [token for token in tokens if token not in string.punctuation]
    tokens = [token for token in tokens if token not in stopwords.words('english')]
    return tokens

# Preprocess each document
processed_docs = [preprocess(doc) for doc in docs]

```

## Creating a Dictionary and Corpus
```{python py create dictionary and corpus}

# Create a dictionary representation of the documents
dictionary = corpora.Dictionary(processed_docs)

# Filter out extremes to remove noise and infrequent words
dictionary.filter_extremes(no_below=5, no_above=0.5, keep_n=100000)

# Convert the dictionary to a bag-of-words corpus
corpus = [dictionary.doc2bow(doc) for doc in processed_docs]

```

## Applying LDA Topic Modeling
```{python py topic modeling}

# Apply LDA model
lda_model = models.LdaModel(corpus, num_topics=10, id2word=dictionary, passes=10)

# Print the topics
#for index, topic in lda_model.print_topics():
#    print(f"Topic: {index} \nWords: {topic}\n")

```

## Visualizing the Topics
```{python py topic visualization}

import pyLDAvis
import pyLDAvis.gensim_models as gensimvis

# Prepare visualization of the topics
vis = gensimvis.prepare(lda_model, corpus, dictionary)

# Save the visualization to an HTML file
pyLDAvis.save_html(vis, 'lda_visualization.html')

```